% !TEX enableShellEscape = yes
% (The above line makes atom-latex compile with -shell-escape for
% minted, and is just ignored by other systems.)

%====================================================
% Please read the README.txt file first.
%====================================================

% Change the "masc" command to your degree title
\documentclass[masc,oneside]{ubcthesis}

% Looser page layout (increase margins slightly)
\usepackage[margin=1.25in]{geometry}

% Some packages that are necessary and that may be useful. Be careful about removing any as it may cause some
% formatting errors.
\usepackage[unicode=true,
  linktocpage,
  linkbordercolor={0.5 0.5 1},
  citebordercolor={0.5 1 0.5},
linkcolor=blue,
  hypertexnames=false]{hyperref}
\usepackage{amsmath, amsfonts, amssymb}
\usepackage{graphicx}
\usepackage{afterpage}
\usepackage{float}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{pdflscape}
\usepackage{rotating}
\usepackage[numbers,sort&compress]{natbib}
\usepackage{psfrag}
\usepackage{makecell}
\usepackage{parskip}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{enumitem}
\usepackage{minted}
\usepackage[normalem]{ulem}
\usepackage{float}
% This package is for generating the dummy text you'll see on some pages.
\usepackage{lipsum}

\usepackage{palatino}  % Latin Modern, the modern version of Computer Modern

% New command for the Examining Committee Page
\newcommand{\committeeMember}[4]{#1, #2, #3\\\textit{#4}}

\newcolumntype{Y}{>{\centering\arraybackslash}X}

%================================================
% THESE COMMANDS ARE REQUIRED
\def \theAuthor {Randy Zhu}
\def \theTitle {ExploTest, a Python Unit Test Carving Tool}
% \def \theSubtitle {Thesis Subtitle, if applicable}
\def \theDegree {Bachelor of Science}
\def \theProgramTitle {Honours Computer Science, Option in Software Engineering}
\def \theCopyrightyear {2026}
\def \theSubmitdate {April 2026}
%===============================================

% These commands are optional.  The defaults are shown.  You only
% need to include them if you need a different value
\institution{The University Of British Columbia}

% If you are at the Okanagan campus, then you should specify these
% instead.
\faculty{The Faculty of Science}
%\institutionaddress{Okanagan}
% \faculty{The Faculty of Graduate and Postdoctoral Studies}
\institutionaddress{Vancouver}

% You can issue as many of these as you have...
% \previousdegree{B.ASc., The University of British Columbia, 2021}

% The ubcthesis class has no built-in bachelor option, so we override the
% class-set degree title (which would otherwise read "Master of Applied Science").
\degreetitle{Bachelor of Science}

% These commands pass the above definitions to the class file to generate the title page
\title{\theTitle}
% \subtitle{\theSubtitle}
\author{\theAuthor}
\copyrightyear{\theCopyrightyear}
\submitdate{\theSubmitdate}
% \submitdate{\monthname\ \number\year} % The "\ " is required after
% \monthname to prevent the
% command from eating the space.
\program{\theProgramTitle}

% These commands make the page numbering consistent with G+PS
\renewcommand\thepart         {\Roman{part}}
\renewcommand\thechapter      {\arabic{chapter}}
\renewcommand\thesection      {\thechapter.\arabic{section}}
\renewcommand\thesubsection   {\thesection.\arabic{subsection}}
\renewcommand\thesubsubsection{\thesubsection.\arabic{subsubsection}}
\renewcommand\theparagraph    {\thesubsubsection.\arabic{paragraph}}
\renewcommand\thesubparagraph {\theparagraph.\arabic{subparagraph}}

\setcounter{tocdepth}{2}
\setcounter{secnumdepth}{2}

% Here is an example of a "Program" environment defined with the
% "float" package.  The list of programs will be stored in the file
% ubcsample.lop and the numbering will start with the chapter
% number.  The style will be "ruled".
\floatstyle{ruled}
\newfloat{Program}{htbp}{lop}[chapter]

% Used only for sending to reviewers
% \usepackage{setspace}
% \onehalfspacing

\begin{document}

%% This starts numbering in Roman numerals as required for the thesis style and is mandatory.
\frontmatter

% This next section is for the admin pages. These commented lines exist to tell you what you need

%%% The order of the following components should be preserved.  The order
%%% listed here is the order currently required by FoGS:        \\
%%% Title (Mandatory)                                           \\
%%% Committee Page (Mandatory)                                  \\
%%% Preface (Manditory if any collaborator contributions)       \\
%%% Abstract (Mandatory)                                        \\
%%% List of Contents, Tables, Figures, etc. (As appropriate)    \\
%%% Acknowledgements (Optional)                                 \\
%%% Dedication (Optional)                                       \\

% Now
\maketitle                      %% Mandatory
\include{admin_chapters/committee_page/committee_page} %% Very mandatory

\begin{abstract}                %% Mandatory -  maximum 350 words
  Unit tests are costly to write. Both the inputs and the oracles take effort, and researchers have built techniques to
  generate them automatically. One such technique, unit test carving, extracts unit tests from system tests.
  ExploTest, the tool studied in this thesis, carves tests for Python by serializing the arguments and return
  values of the code-under-test during a system test run. We ask whether this cheap, state-based approach produces
  tests that catch bugs, and whether carving is worth the cost, in both performance and developer overhead in using the tool.

  We evaluate ExploTest on 22 bugs across 12 popular open-source Python repositories drawn from BugsInPy \cite{widyasari2020bugsinpy} and QuAC \cite{wu2024quac}.
  Of these 22 bugs, only 2 are exposed by a generated test. This is not primarily a capture failure: the tool correctly exercises the function-under-test
  for 14 of the 22 bugs. The drop-off happens at oracle generation, and this work does not resolve its root cause. Exercise-side failures that do occur
  concentrate on a single cause: unserializable operating-system resources, which our serialization library, dill, cannot capture. Carved unit test suites
  are faster than their system tests in 33\% of cases, with speedup scaling strongly with system test runtime ($R^2 = 0.849, p < 0.001$). Carving
  imposes a $1.43 \times$ geometric-mean overhead on the system-test run itself.

  A proposed usage procedure for ExploTest aligns with five of seven steps in Zeller's TRAFFIC systematic debugging procedure \cite{zeller2009programs},
  suggesting carving integrates with debugging work developers already perform.
  We conclude that state-based carving is viable as a byproduct of debugging, capturing the concrete bug-triggering inputs surfaced during
  bug localization as the skeleton of a unit test, but is not yet a standalone bug-detection tool.
  The more pressing direction is oracle generation, since tests that exercise the code-under-test correctly still fail to expose the bug.
  Oracles should move beyond return-value comparison to capture post-execution state, through program-state instrumentation like Pythoscope \cite{wiatkowski2011pythoscope},
  or LLM-guided property-based testing.
  A secondary direction that addresses the exercise-side failures is action-based carving,
  which reconstructs objects by replaying their construction sequences rather than serializing state,
  extending the tool's reach to operating-system-level resources that cannot be serialized.


\end{abstract}

% These statements generate the rest of the admin pages
\include{admin_chapters/lay_summary/lay_summary}    %% Mandatory
\include{admin_chapters/preface/preface}            %% Mandatory
\tableofcontents                %% Mandatory
\listoftables                   %% Mandatory if thesis has tables
\listoffigures                  %% Mandatory if thesis has figures
\include{admin_chapters/list_of_abbreviations/list_of_abbreviations}    %% Optional
\include{admin_chapters/list_of_symbols/list_of_symbols}                %% Optional
\include{admin_chapters/acknowledgements/acknowledgements}              %% Optional
\include{admin_chapters/dedication/dedication} %% Optional

% Switch the numbering to arabic
\mainmatter
% THIS IS THE MAIN BODY OF YOUR THESIS.
% Add new chapters as you see fit. It will be useful to retain the structure already defined in the
% body_chapters folder.You will have to manually
% add and remove \include statements like these as you add and remove chapters.

\include{body_chapters/chap_1/chap_1}
\include{body_chapters/chap_2/chap_2}
\include{body_chapters/chap_3/chap_3}
\include{body_chapters/chap_4/chap_4}
\include{body_chapters/chap_5/chap_5}

% Insert the bibliography
\bibliographystyle{ACM-Reference-Format} % We choose the "unsrt" reference style
\bibliography{sample_bib} % Entries are in the sample_bib.bib file

% ========================================
% APPENDICES
% ========================================
\appendix
% Add as many of the following statements as you need.
\include{body_chapters/chap_a1/chap_a1}
\include{body_chapters/chap_a2/chap_a2}
\include{body_chapters/chap_a3/chap_a3}
\include{body_chapters/chap_a4/chap_a4}
\include{body_chapters/chap_a5/chap_a5}

\end{document}
