\chapter{Background}


\section{Unit Test Carving}

Unit test carving, originally called unit test factoring, was introduced by Saff et al \cite{saff2005automatic}. Test carving is
described to be ``a technique for
automatically creating fast, focused unit tests from slow system-wide tests; each new unit test exercises
only a subset of the functionality
exercised by the system tests" \cite*{saff2004mock, saff2005automatic}. Test carving techniques typically
involve record-and-replay (also called
carve-and-replay \cite*{elbaum2006carving, elbaum2008carving}), recording function executions, serializing
the inputs and outputs of the function,
then invoking the function again with said captured inputs.

Elbaum et al. \cite{elbaum2006carving} coined the term unit test carving. Abstractly, carving a unit
test for a given method $m$ involves capturing the program state immediately before and after $m$'s
execution, and the pair of these pre- and post-execution states constitutes a unit test for $m$. As
the codebase of a program evolves, $m$ may change over time.

We adopt the more granular formalism of Kwiatkowski \cite{l2011capture}, who models a function-under-test
call as a transformation $f : (A, S) \rightarrow (R, E, S')$, where $A$ is the arguments to the function,
$S$ is the starting external state, $R$ is the return value, $E$ are the exceptions thrown, and $S'$ is
the new external state.

The execution of a carved unit test must first restore $(A, S)$, execute $m$, then compare the observed
$(R, E, S')$ to the captured $(R, E, S')$. In common software engineering parlance, the execution of a
carved unit test is as follows:

\begin{itemize}
  \item Restoring $(A, S) \rightarrow$ test setup; typically found in the fixtures or the
        \verb|@BeforeEach| part of a unit test.
  \item Executing $m \rightarrow$ running the function-under-test
  \item Comparing observed $(R, E, S')$ to captured $(R, E, S') \rightarrow$ the test oracle; typically
        assertions like \verb|assertEquals|.
\end{itemize}

% There are two approaches in capturing $(A, S)$, explains Elbaum et al.: state-based and action-based.

State-based and action-based carving are two approaches to unit test carving, explains Elbaum et al.

State-based capture involves saving $(A, S)$, typically serializing it to disk, then deserializing to
restore it. This comes with certain challenges which we will describe later in this thesis. The majority
of unit test carving tools use state-based capture.

Action-based capture involves saving the instructions that produced $(A, S)$, and later replaying those
instructions to re-create it. This resolves many of the challenges of state-based capture discussed in
section 2.1.1, but is harder to do in practice.

%ExploTest has a limited
% version of action-based capture in its
% argument reconstruction mode.

\subsection{Challenges in State-based Capture}

These challenges all concern the capture and restoration of $(A, S)$: either it cannot be serialized
faithfully, or the serialized form becomes unmaintainable.

\subsubsection{Evolution and Serialization}
Serialization involves storing the state of the program, $(A, S)$, to disk \cite{isocppserialization}. In turn, when we
want to restore the state of the
program, we have to load that state from disk. This state is entirely unaware of changes to the source code,
which means any changes to the structure
of the object in the source code at a later time is not reflected in the serialized format. In turn, this
makes the unit tests more brittle to
breaking from code evolution.

\subsubsection{Serialization Opaqueness}
Unlike the code to create objects in source code, reading objects from disk does not at all explain the
process that got the object to that
serialized state. This makes it much more confusing for maintainers of source code that utilizes serialized
state in unit tests to change the unit
tests as code evolves, worsening the evolution issue.
MicroTestCarver by Deljouyi and Zaidman \cite{deljouyi2023generating} tries to remedy this issue in their
work by serializing to XML, a
human-editable file format using the XStream library, however, there is no similar library that produces any
kind of human-readable serialization
output in Python.

\subsubsection{Non-serializable Objects}

There are objects that cannot be serialized, which typically have ill-defined representations after
deserialization. Currently, the most robust
Python serialization library which we elected to use, dill \cite{dilldocs} does not support \verb|generator|,
\verb|traceback|, and code frame objects. Code frame
objects cannot be captured by state-based capture, as the creation of a code frame object entails the
creation of native code objects which cannot be
done without some action-based capture. Similarly, network sockets and active network connections do not have
a meaningful value when deserialized,
as the server they connect has likely since closed the connection. Finally, objects which explicitly reject
serialization by disabling
\verb|__reduce__| cannot be serialized. This family of objects necessitates the use of action-based capture.

\subsection{Challenges in Action-based Capture}

\subsubsection{The Object Creation Problem}
The Object Creation Problem, defined by Bach et al. \cite{bach2020determining}, is the problem of finding the
correct sequences of method calls to
construct an object. In Python, creating objects is not always as simple as calling the \verb|__init__|
constructor, instead, objects rely on
hierarchies and overridden \verb|__new__|, or other classes like in the factory design pattern.

\subsubsection{Arbitrary Object Modification}
Explained by Politz et al. \cite{politz2013python}, the semantics of Python allow for arbitrary modification
to object attributes at any time in the
programs. The shape of an object, which we define to be the number and names of fields of an object can
change at any point in a Python program.

While this is a problem in unmanaged languages like C++, and the authors of \cite{bach2020determining}
discuss this issue, they consider it out of
scope of the object creation problem, as it is unidiomatic in C++, and often leads to undefined behaviour.
Unfortunately, this technique of arbitrary
modification of objects at any point is often used by Python programmers.

This presents challenges to action-based capture, as existing action-based capture unit test carving tools
like ARTISAN \cite{gambi2023action} assume
that the shape of an object does not change.

\section{Prior Work in Test Carving for Python}
\subsection{Pythoscope}

Pythoscope, by Kwiatkowski and Clements-Tiberio \cite*{pythoscopewikidot,pythoscopegithub,l2011capture} is an
open source Python 2 unit test
carving tool, created around 2008.

Pythoscope performs both static and dynamic analysis of the system test run.

The static inspection pass parses source code into abstract syntax trees, then the tool identifies classes, functions,
and module structure. The dynamic inspector achieves per-bytecode tracing using CPython's \verb|sys.settrace|
combined with Batchelder's line-number-table rewriting hack \cite{batchelder2008bytecode}, which makes every
bytecode instruction appear as its own line so that the \verb|'line'| event fires for each one.
Using these two techniques, Pythoscope records every function call, attribute access, and global variable read write, which
gets stored into a project tree.

Using the $f : (A, S) \rightarrow (R, E, S')$ formalism introduced in section 2.1, Pythoscope captures all
of $A$, $S$, $R$, $E$, and $S'$.

External state, $S$ and $S'$ are handled in a rules-based manner. Different effects are classified using a rules-based system that
maps specific bytecode instructions and C function calls to known side-effect types (e.g., the instruction \verb|STORE GLOBAL| indicates
global variable rebinding, calls to \verb|list.append| indicates mutation of a global variable).

The handling of function arguments $A$ and function return values $R$ are remarkably similar to ExploTest. Pythoscope instruments
the bytecode for \verb|CALL_FUNCTION| instructions, and then serializes the function arguments and results, just like ExploTest.
Pythoscope handles serialization failure more gracefully than ExploTest, however, when serialization of a function argument fails,
Pythoscope emits a test stub with a TODO comment, allowing the developer to complete the test manually. Like ExploTest, the
serialization-based approach fails on objects that hold operating system resources such as sockets or file descriptors.

Pythoscope has an action-based oracle generation approach, since it instruments the entire program execution. It records external state
throughout the function, and creates oracles involving global state by analyzing the various different global states during
program execution.

Pythoscope has been abandoned, its last commit was on 2016-08-23, as of 2026-03-26. It only supports Python 2.7.

\subsection{Auger}
Auger, created by Laffra, et al., \cite{augerblogspot} is another Python 3 unit test carving tool. It functions more similarly to
the previous iteration of ExploTest. The tool works by using the \verb|sys.settrace| API, which intercepts every call and return during
program execution. It takes note of candidate function-under-tests, and if the function-under-test belongs to a class, it performs
simple, action based capture by re-constructing the call the \verb|__init__|. All serialization is done via \verb|__repr__|.
Since \verb|__repr__| does not produce valid Python for many object types (e.g., sockets, file handles), Auger can silently emit
invalid test files without warning.

Auger has been abandoned, its last commit was on 2019-02-21, as of 2026-03-26. It only supports Python 3.5.

% \subsection{Fuzzing Book Carver}
% \cite{fuzzingbook2023:Carver}

\section{ExploTest Architecture}

ExploTest is a unit test carving tool, originally conceived by Zihao Huang under the supervision of Dr. Caroline Lemieux and substantially rewritten for this thesis. A full attribution appears in the preface.


ExploTest can be used in an exploratory manner for exploratory testing or for creation regression test suites
from system test runs.
As defined by Kaner \cite{kaner2008tutorial}, exploratory testing is an unscripted technique where the software designer verifies the program's
correctness using their own intuition and domain knowledge to uncover defects and create test oracles. For this thesis, it primarily
focuses on a tool to create regression test suites.

\subsection{ExploTest Implementation}
The user of ExploTest marks a carve candidate using the \verb|@explore| annotation.

\subsubsection{Argument building}
ExploTest performs static and dynamic inspection of the test run, and has many similarities to Pythoscope. Its static pass first builds the
test AST using the \verb|ast| library built into Python.

The static pass collects all the dependencies of the function by analyzing the imports, and then
sets up the required Python fixtures for function arguments. Then, the tool performs a dynamic analysis using the \verb|@explore| annotation,
which binds all the arguments to the function-under-test, (e.g., the function \verb|f(x: int, y: int, z: int)|, called with \verb|f(3,4,5)|
would have bindings \verb|{x: 3, y: 4, z: 5}|), and, depending on the mode, either serializes them using the \verb|dill| package, or applies the ARR technique defined below.
Once serializing the arguments is complete,
the assertion generation phase of the tool begins.


\subsubsection{Assertion building}
Assertions are generated using a derivative of AutoAssert by Zamprogno et al. \cite{zamprogno2022dynamic}. We choose AutoAssert for its
human-in-the-loop factor, since we expect the user to inspect and edit the unit test after generation.

ExploTest only generates assertions on the
return value, $R$, not the external state after the function executes, $S'$. This means if the function modifies \verb|self|, for instance,
the assertion generator is unable to capture that behaviour.

AutoAssert is rules based, and requires the function-under-test be run twice. Given
two execution results $R_1$ and $R_2$, AutoAssert tries to apply the following ruleset.

\begin{itemize}
  \item Are $R_1$ and $R_2$ both primitives representable with a Literal (e.g., \verb|str|, \verb|int|, \verb|float|, \verb|None|, \verb|bool|)? $\rightarrow$
        Create a comparison against the literal.
  \item Are $R_1$ and $R_2$ equal under their built-in \verb|__eq__| method? $\rightarrow$ Perform an equality assertion with $R_1$.
  \item Does the function have a custom \verb|__repr__| method, and are reprs of $R_1$ and $R_2$ equal? $\rightarrow$ Assert against the \verb|__repr__|s of $R_1$.
  \item Does $R_1$ and $R_2$ have a length and are they equal? $\rightarrow$ Create an assertion on the length.
  \item Are the types of $R_1$ and $R_2$ equal under \verb|type()|? $\rightarrow$ Assert against the name of the type (e.g., \verb|type(return_value).__name__ == "str"|)
  \item If none of the criteria are met, do not create an assertion.
\end{itemize}

\subsubsection{Argument-reconstruction-reconstructor (ARR) technique}

One concern with serialization is readability and how well serialized arguments lend themselves to source code evolution.
The argument-reconstruction-reconstructor (ARR) technique tries to simulate some of the benefits of action-based carving statically (it does not
perform any dynamic code instrumentation).

Rather than serializing an object to bytes, the ARR inspects the argument's attributes at runtime using \verb|inspect.getmembers()|, and
generates code that reconstructs a clone. It bypasses the constructor by instead creating a skeleton with \verb|__new__|, and then populating
each attribute with \verb|setattr|. For example, given an instance of \verb|MyClass| with attribute \verb|name = "hello"|, and \verb|count = 67|, the ARR generates:

\begin{verbatim}
  arred = module.MyClass.new()
  setattr(arred, 'name', 'hello')
  setattr(arred, 'count', 67)
\end{verbatim}

Before attempting the techniques, ARR performs a reachability check. It traverses the object's attribute graph via breadth-first search
to verify every attribute can be meaningfully reconstructed. If an object cannot be reconstructed, it falls back to pickling. Self-referential
objects are handled greedily, and use a technique similar to Landin's Knot to handle self-referential objects \cite{koronkevich2025one}.

The ARR approach has an advantage over Auger's constructor-replay technique, it does not require observing how the object was originally
constructed. Objects built through the Factory or Builder pattern can all be reconstructed. However, the trade-off is that it still does not
exercise \verb|__init__|, which may set up required external state to the function under test, or enforce invariants.

\subsubsection{Comparison with Pythoscope}
ExploTest and Pythoscope are independent state-based carving tools for Python, separated by roughly fifteen years and one major language version.
Table \ref{tab:pythoscope-explotest} summarizes their differences. The central contrast is one of implementation surface.
Pythoscope's tracer depends on per-bytecode tracing, an import-hook-driven code rewriter, and \verb|ctypes| value-stack inspection to capture the full
$f : (A, S) \rightarrow (R, E, S')$ formalism, while ExploTest captures the narrower subset $(A, R, E)$ using a decorator
and a static AST pass. On unserializable arguments,
Pythoscope degrades to a test stub with a TODO comment; ExploTest terminates.
Whether ExploTest's simpler design and narrower capture surface meaningfully impacts test quality is a question this thesis addresses empirically.

\begin{table}[h]
  \centering
  \caption{Comparison of Pythoscope and ExploTest design choices.}
  \label{tab:pythoscope-explotest}
  \begin{tabularx}{\linewidth}{l X X}
    \toprule
     & \textbf{Pythoscope}                                    & \textbf{ExploTest} \\
    \midrule
    Formalism coverage
     & $A, S, R, E, S'$
     & $A, R, E$                                                                   \\
    \addlinespace
    Capture mechanism
     & Bytecode tracing
     & \verb|@explore| decorator + static AST pass                                 \\
    \addlinespace
    Argument serialization
     & Stack snapshot with rules-based side-effect classifier
     & \verb|dill| (pickle-mode) or ARR                                            \\
    \addlinespace
    Oracle generation
     & Rules-based over captured $S'$ deltas
     & AutoAssert over $R$                                                         \\
    \bottomrule
  \end{tabularx}
\end{table}



